---
title: "All you need is Rmarkdown!"
output: 
  bookdown::html_document2:
    keep_md: True
    toc: no
---

# Introduction

- 1) Diversity of options to write research papers or academic slides.
    - Word, Latex, PowerPoint, Beamer

- 2) Separation between manuscript and code
    - Increase the risk of typing errors.
    - Increase the amount of work spent in non-static attributes (i.e. figures and tables in word).
    - The reproducibility of the paper become more difficult.

- 3) A solution: `Rmarkdown`
    - Execute R, Python or Stata in one place to produce your statistical analysis.
    - Use $\LaTeX$ for equations and more!
    - Markdown syntax for the body of the text.
    - Configure the document with a `YAML` header.

- 4) Focus of the post
  - Pdf documents.
  - Mention the other types (HTML, Slides, Dashboards)

# Executing code within an Rmarkdown Document.
```{r Loading Data}
library(datasets)
data(iris)

mean_Sepal.Lenght <- mean(iris$Sepal.Length)
```

I usually run a `Setup` chunk for all the Rmarkdown file. In all my projects, I include a [settings.R](https://gist.github.com/ignacioriveros1/797bc12b7eb161caba88c43cd7e6fe78) script where I load all the packages for the project and the local paths of the users that works on them.

```{r Setup}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = F,
  error = F,
  tidy = T,
  tidy.opts = list(width.cutoff = 50),
  collapse = T,
  warning = FALSE,
  error = FALSE,
  message = FALSE,
  comment = "",
  fig.cap = " "
)
# These options are tuned for manuscript/presentation.
# They basically run R in the background except for spitting out figures/tables
# They also cache results so that you don't have to run your whole analysis every time you fix a typo

# source('settings.R', echo = FALSE)
```


# Executing Inline Code
To execute inline code you have to use the following: \` r yourcode  \`.

For example, using the chunk of above, I can introduce the mean length of Sepal in my text, which is `r mean_Sepal.Lenght`, using ` r mean_Sepal.Lenght`. Alternatively, I could calculate the same value (`r mean(iris$Sepal.Length)`) using the function in the inline code directly (` r mean(iris$Sepal.Length)`). Also, If I don't like the number of decimals, I can use ` r round(mean_Sepal.Lenght)` function to convert `r mean_Sepal.Lenght` into `r round(mean_Sepal.Lenght,2)`.

# Tables 

# Figures
```{r FigureExample, results='asis', fig.cap="Quadratic Function"}
x <- 1:100
y <- (1:100)^2

plot(x, y)

```

Here I can reference Figure \@ref(fig:FigureExample)


# LaTeX
$$E = mc^2$$

# Markdown

# Bibliography

# Configuring the document with YAML

# My manuscript template.

# Final Thoughts

# References and Further reading
- [R markdown from RStudio](https://rmarkdown.rstudio.com/lesson-1.html)
- [R markdown from RStudio: Cheatsheet](https://rmarkdown.rstudio.com/lesson-15.html)
- [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/)
- [RMarkdown for Scientists](https://rmd4sci.njtierney.com)

