---
title: "All you need is Rmarkdown!"
output: 
  bookdown::html_document2:
    keep_md: True
    toc: no
---

# Introduction

- 1) Diversity of options to write research papers or academic slides.
    - Word, Latex, PowerPoint, Beamer

- 2) Separation between manuscript and code
    - Increase the risk of typing errors.
    - Increase the amount of work spent in non-static attributes (i.e. figures and tables in word).
    - The reproducibility of the paper become more difficult.

- 3) A solution: `Rmarkdown`
    - Execute R, Python or Stata in one place to produce your statistical analysis.
    - Use $\LaTeX$ for equations and more!
    - Markdown syntax for the body of the text.
    - Configure the document with a `YAML` header.

- 4) Focus of the post
  - Pdf documents.
  - Mention the other types (HTML, Slides, Dashboards)
  - Emphasize that this post is not an exhaustive revision of all the capabilities of Rmarkdown.

# Creating a Rmarkdown document

# Markdown
In my opinion, one of the features that make Rmarkdown a easy to learn tool is Markdown. "*Markdown is a text-to-HTML conversion tool for web writers. Markdown allows you to write using an easy-to-read, easy-to-write plain text format, then convert it to structurally valid XHTML (or HTML).* ([John Gruber](https://daringfireball.net/projects/markdown/))". This language is super easy to learn and it is simple to format. For instance:
```
## Header

*Italic Letters*

**Bold Letters**

* List
* List
* List

1. Numbered List
2. Numbered List
3. Numbered List

`code`

Hello, I'm writing some text in my Rmarkdown document. Markdown is super easy to use!

[Check more tips and tricks](https://www.markdownguide.org/cheat-sheet/)
```

## Header

*Italic Letters*

**Bold Letters**

* List
* List
* List

1. Numbered List
2. Numbered List
3. Numbered List

`code`

Hello, I'm writing in my Rmarkdown document. Markdown is super easy to use!

[Check more tips and tricks](https://www.markdownguide.org/cheat-sheet/)

# Executing code within an Rmarkdown Document.
Run code in Rmarkdown is really simple. You only need to create a chunk of code using `"```"`. Alternatively, if you use RStudio (which is the IDE that I definitely recommend for coding in R) you can use the option "Insert a new code chunk".

```{r My example Chunk that is not going to be executed, eval=FALSE}
print('Hello World')
```

I usually run a `Setup` chunk at the beginning of all my Rmarkdown documents In all my projects, I include a [settings.R](https://gist.github.com/ignacioriveros1/797bc12b7eb161caba88c43cd7e6fe78) script where I load all the packages thet I will use together with the local paths of the users' machines that will collaborate.

```{r Setup, eval=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = F,
  error = F,
  tidy = T,
  tidy.opts = list(width.cutoff = 50),
  collapse = T,
  warning = FALSE,
  error = FALSE,
  message = FALSE,
  comment = "",
  fig.cap = " "
)
# These options are tuned for manuscript/presentation.
# They basically run R in the background except for spitting out figures/tables

# source('settings.R', echo = FALSE)
```

These chunks support the execution of a wide variety of codes. You can execute Python (using `reticulate`), JavaScript, Bash, SQL querys, and also Stata (check [this post](.) to see how).

# Executing Inline Code
To execute inline code you have to use the following: `(r) yourcode` without the parenthesis in the `r`. Lets see a simple example using the `iris` dataset.
```{r Loading Data}
library(datasets)
data(iris)

mean_Sepal.Lenght <- mean(iris$Sepal.Length)
```

```
For example, using the chunk of above, I can introduce the mean length of Sepal in my text, which is `(r) mean_Sepal.Lenght`. Alternatively, I could calculate the same value (`(r) mean(iris$Sepal.Length)`) using the function in the inline code directly. Also, If I don't like the number of decimals, I can use `round` function to convert `(r) mean_Sepal.Lenght` into `(r) round(mean_Sepal.Lenght,2)`.
```

For example, using the chunk of above, I can introduce the mean length of Sepal in my text, which is `r mean_Sepal.Lenght`. Alternatively, I could calculate the same value (`r mean(iris$Sepal.Length)`) using the function in the inline code directly. Also, If I don't like the number of decimals, I can use `round` function to convert `r mean_Sepal.Lenght` into `r round(mean_Sepal.Lenght,2)`.

# Tables
```{css, echo = FALSE}

table, td, th {
  border: none;
  padding-left: 1em;
  padding-right: 1em;
  margin-left: auto;
  margin-right: auto;
  margin-top: 1em;
  margin-bottom: 1em;
}
```

My favorite way of producing tables in Rmarkdown is with `stargazer`. This package supports a wide variety of functionalities, mainly oriented for regressions tables. Check [this post](link_to_the_post) in which I explain how to use it and show some examples. Here, I will show two examples of how to combine `stargazer` with Rmarkdown.

In the first place, let's produce a descriptive statistic tables. We will use `iris` data:
```{r Des Sta, results='asis', warning=FALSE, message=FALSE}
means_array <-
  c(
    mean(iris$Sepal.Length),
    mean(iris$Sepal.Width),
    mean(iris$Petal.Length),
    mean(iris$Petal.Width)
  )
sd_array <-
  c(
    sd(iris$Sepal.Length),
    sd(iris$Sepal.Width),
    sd(iris$Petal.Length),
    sd(iris$Petal.Width)
  )   
p25_array <-
  c(
    quantile(iris$Sepal.Length, 0.25),
    quantile(iris$Sepal.Width, 0.25),
    quantile(iris$Petal.Length, 0.25),
    quantile(iris$Petal.Width, 0.25)
  )   
p75_array <-
  c(
    quantile(iris$Sepal.Length, 0.75),
    quantile(iris$Sepal.Width, 0.75),
    quantile(iris$Petal.Length, 0.75),
    quantile(iris$Petal.Width, 0.75)
  )

cols <- c('Mean', 'SD', 'P25', 'P75')
rows <-
  c('Sepal.Length', 'Sepal.Width', 'Petal.Length', 'Petal.Width')

data_matrix <-
  matrix(
    data = c(means_array, sd_array, p25_array, p75_array),
    nrow = 4,
    ncol = 4,
    dimnames = list(rows, cols)
  )

library(stargazer)
stargazer(data_matrix, type = 'html', title = 'Summary Statistics Iris Data')
```

If you are working with `html` files and you want to modify tables attributes (width, styles, etc), I recommend to use a custom `.css` file to achieve this. You could declare it inside the document, or in an external file. You don't need to be an expert to program this settings, just google it and change the attributes until you are Ok with the results. If you want to go further, and customize your entire `html` check [this link](https://bookdown.org/yihui/rmarkdown-cookbook/html-css.html). For instance, we are using the following `css`:

```
table, td, th {
  border: none;
  padding-left: 1em;
  padding-right: 1em;
  margin-left: auto;
  margin-right: auto;
  margin-top: 1em;
  margin-bottom: 1em;
}

```

My second example, is mainly directed for $\LaTeX$ users. I personally don't like how `stargazer` package handle the notes below the tables, and also do not support an option to resize the table. In this example, I will show an application of how I solved this.

Let's estimate a model and produce an `stargazer` output. Some important considerations here is that we only want the code within the `tabular` environment. To get this, we need to specify the options `header` and `float` to `FALSE`. Also, as we don't want the notes so, we will leave it empty.
```{r stargazer 2, warning=FALSE, message=FALSE}
# 1) Estimate a linear model
model <- lm(Sepal.Length ~ Sepal.Width + Petal.Length + Petal.Width, data=iris)

# 2) Capture the stargazer output in an object.
out <-
  capture.output(
    stargazer(
      model,
      type = "latex",
      float = FALSE,
      header = FALSE,
      notes.label = '',
      notes = '',
      notes.append = FALSE,
      notes.align = 'l'
    )
  )

# 3) Collapse the array in a string
tex_table <- paste(out, collapse = "\n")

# 4) Define the settings of my table (title, labels, notes)
caption <- 'My title'
label <- 'tab:my_table'
notes <- 'Own elaboration based on iris data.'
```

After we capture the output of stargazer in an object and collapse it in a string, we can use this objects together with inline code execution to fill a custom `.tex` table within our document. The $\LaTeX$ table will look like this:

```
{=latex}
\begin{table}[H]
\centering
\caption{`(r) caption`}
\label{`(r) label`}
\resizebox{\textwidth}{!}{
`(r) tex_table`
}
\begin{minipage}{\textwidth}
\tiny \textbf{Notes}: `(r) notes` \par
\end{minipage}
\end{table}
```

```
{=latex}
\begin{table}[H]
\centering
\caption{`r caption`}
\label{`r label`}
\resizebox{\textwidth}{!}{
`r tex_table`
}
\begin{minipage}{\textwidth}
\tiny \textbf{Notes}: `r notes` \par
\end{minipage}
\end{table}
```

I want to point to important observations. In the first place, you must declare `{=latex}` in the chuck's header. This will allow Rmarkdown and the compiler that all the code that is within that chunk must be interpreted as `LaTex`, This will save you a lot of problems while compiling the document to pdf. In the second place, remember that it is necessary to include the package `float` in your document if you want to specify the option `[H]` in the `table` environment. I will cover this topic later!

# Figures
To include a figure, just create it using `plot` or `ggplot`.

```{r FigureExample, fig.cap="Quadratic Function"}
x <- 1:100
y <- (1:100)^2

plot(x, y)

```

If you don't want the code chunk to appear in the document, include the chunk with the option `echo=FALSE`. Also, some functions will print some warnings and messages. If you want to turn them off, include `warning=FALSE, message=FALSE` in the preamble of your code chunk.
```
{r Figure2, fig.cap="My amazing figure", echo=FALSE, warning=FALSE}
# The code of this plot was extracted from https://www.publichealth.columbia.edu/sites/default/files/media/fdawg_ggplot2.html
library(ggplot2)
density2 <- ggplot(data=iris, aes(x=Sepal.Width, fill=Species))
density2 + geom_density(stat="density", alpha=I(0.2)) +
  xlab("Sepal Width") +  ylab("Density") + ggtitle("Histogram & Density Curve of Sepal Width")
```

```{r Figure2, fig.cap="My amazing figure", echo=FALSE, warning=FALSE}
# Plot from https://www.publichealth.columbia.edu/sites/default/files/media/fdawg_ggplot2.html
library(ggplot2)
density2 <- ggplot(data=iris, aes(x=Sepal.Width, fill=Species))
density2 + geom_density(stat="density", alpha=I(0.2)) +
  xlab("Sepal Width") +  ylab("Density")
```

```
Using `bookdown::html_document2`, I can reference Figure \@ref(fig:FigureExample) and \@ref(fig:Figure2). To make this work, it is important to label your figure chunks without spaces.
```
Using `bookdown::html_document2`, I can reference Figure \@ref(fig:FigureExample) and \@ref(fig:Figure2). To make this work, it is important to **label your figure chunks without spaces or special characters**.


# LaTeX
```
Rmarkdown comes with $\LaTeX$ incorporated. You can produce environments, equations, and use most of LaTeX cabalities inside your notebooks. Let's see some examples:

$$E = mc^2$$
```

Rmarkdown comes with $\LaTeX$ incorporated. You can produce environments, equations, and most of LaTeX capabiliteies inside your notebooks. Let's see some examples:

$$E = mc^2$$

You also, can number equations using `equation` environment together with a label.
```
\begin{equation}
\hat{\beta} = (X'X)^{-1}X'Y (\#eq:eq1)
\end{equation} 
 
The model \@ref(eq:eq1) is the estimation by OLS.
```
\begin{equation}
\hat{\beta} = (X'X)^{-1}X'Y (\#eq:eq1)
\end{equation} 
 
The model \@ref(eq:eq1) is the estimation by OLS.

You also can include `tikz` graphs inside your Rmarkdown documents. Check this [SO question](https://stackoverflow.com/questions/27880563/tikz-in-r-markdown) to see more information.

```
\begin{tikzpicture}
\begin{axis}[xmax=9,ymax=9, samples=50]
  \addplot[blue, ultra thick] (x,x*x);
  \addplot[red,  ultra thick] (x*x,x);
\end{axis}
\end{tikzpicture}

# Code from https://stackoverflow.com/questions/27880563/tikz-in-r-markdown
```

# Configuring the document with YAML
Another thing that I love from Rmarkdown is that the global configuration of the document is with `YAML`. According to [Wikipedia](https://en.wikipedia.org/wiki/YAML), ""*YAML (a recursive acronym for "YAML Ain't Markup Language") is a human-readable data-serialization language. It is commonly used for configuration files and in applications where data is being stored or transmitted*".

This way of configuring a document is far more easier than $\LaTeX$ preambles (not recommended for humans), mainly because YAML is readable. For example, the header of this Rmarkdown looks like this:

```{yaml}
title: "All you need is Rmarkdown!"
output: 
  bookdown::html_document2:
    keep_md: True
    toc: no
```

For example, if I would like to add a Table of Contents and not keep the Markdown file as output of the Rmarkdown, the header will look like:
```{yaml}
title: "All you need is Rmarkdown!"
output: 
  bookdown::html_document2:
    keep_md: False
    toc: yes
```

In addition, this is the place to include a [bibliography](https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html) for your document. Rmarkdown supports `json`, `bibtex`, and much other formats!
```{yaml}
title: "All you need is Rmarkdown!"
output: 
  bookdown::html_document2:
    keep_md: False
    toc: yes
bibliography: bibliography.extension
```

For more configurations, I suggest to check [this guide](https://bookdown.org/yihui/rmarkdown/documents.html).

# My manuscript template.

# Final Thoughts

# References and Further reading
- [R Markdown from RStudio](https://rmarkdown.rstudio.com/lesson-1.html)
- [R Markdown from RStudio: Cheatsheet](https://rmarkdown.rstudio.com/lesson-15.html)
- [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/)
- [R Markdown for Scientists](https://rmd4sci.njtierney.com)
- [Data Visualization with ggplot2](https://www.publichealth.columbia.edu/sites/default/files/media/fdawg_ggplot2.html)
- [TikZ in R Markdown](https://stackoverflow.com/questions/27880563/tikz-in-r-markdown)
- [Markdown Cheat Sheet](https://www.markdownguide.org/cheat-sheet/)
- [YAML](https://en.wikipedia.org/wiki/YAML)
- [Bibliographies and Citations](https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html)